visualizemi: Visualization, Effect Size, and Replication of Measurement Invariance for Registered Reports

### 2024-03-09 manuscript.Rmd
finalizing paper

Therefore, purpose of this manuscript is describe an *R* package, `visualizemi`, that provides functionality to calculate multi-group confirmatory factor analysis, partial invariance tests, visualizations of the size of non-invariance, and potential effect sizes for overall models and individual parameters. No known visualization techniques have been proposed for measurement invariance. By creating panel visualizations, we can supplement a researchers ability to judge the strength of the non-invariance differences and effect size for each item. The proposed effect sizes demonstrate the likelihood of replication with a similar sample as compared to a randomly assigned group model, thus, illustrating what type of measurement one might expect to find, and how different that is from random chance. Within this technique, the individual parameter effect sizes can calculated: both the group differences within a model as compared to random and the likelihood of a parameter replication compared to random groups. Coupled with other indicators (i.e., fit indices differences, $d_{MACS}$), we can move toward a better understanding of how much measurement non-invariance is meaningful. This tutorial and package will help researchers plan future studies and aid in the ability to estimate a smallest effect of interest for measurement invariance studies, rather than relying on fit indices and rules of thumb alone.

@@

1.  Learn how to use *visualizemi* to analyze multi-group confirmatory factor analysis, examine partial invariance, and create visualizations of parameters.
2.  Learn how to estimate the potential replication of multi-group models and their parameters using bootstrapping compared to a random group model.
3.  Be able to calculate and interpret effect sizes for model and parameter replication, as well as parameter group differences.
4.  Understand the impact of measurement variability on replication and generalizability.

@@

1)  `complete`: The output from this model can be found in Figure \@ref(fig:invariant-pic). On the left hand side, the item invariance is plotted, and on the right hand side, the latent mean distributions for the two groups are plotted. In the item invariance sub-plot, the visualization includes all three components traditionally seen in MGCFA testing steps: loadings, intercepts, and residuals. Each visualization element was designed to match the traditional visualization for that type of output. All parameter estimates are plotted on the unstandardized estimates and their confidence interval based on the standard error of the estimate. All plots are made with *ggplot2* and *cowplot* [@wilke2020].

@@

2)  `intercept`: Only the left hand side of the complete plot designed to represent intercepts and factor loadings. Factor loadings represent the slope of the regression equation for the latent variable predicting the scores on the observed variable ($\hat{Y} \sim b_0 + b_1X + \epsilon$). The y-axis indicates the observed variable scores, and here, the plot includes the entire range of the scale of the data for item four. The coefficient ($b_1$) for group 1 was `r apa_num(results.invariant$model_coef %>% filter(model == "Configural") %>% filter(grepl("q4", term)) %>% filter(op == "=~") %>% slice_head() %>% pull(estimate))`, while the coefficient for group 2 was `r apa_num(results.invariant$model_coef %>% filter(model == "Configural") %>% filter(grepl("q4", term)) %>% filter(op == "=~") %>% slice_tail() %>% pull(estimate))`. The ribbon bands around the plotted slopes indicate the confidence interval for that estimate. In this plot, while the coefficients for each group are not literally equal, the overlapping and parallel slope bands indicate they are not different practically.

@@

3)  `mean`: The right hand side of the complete plot graphing the latent variable means and density from the data. The latent variable is shown on the x-axis using standardized values (i.e., *z*-scores) where -1 indicates one standard deviation below the mean for the latent variable, 0 indicates the mean for the latent variable and so on. The lines indicate the means of the latent variables from the simulated dataset. Group labels are represented in the figure caption on the bottom. Group 1 is usually the group that is alphabetically first in the data set or whichever group is the first that appears when using the `levels()` command.

@@

```{r boot-rr, echo = T, eval = F, size = "small"}
boot.model.invariant <- 
  bootstrap_model(
    # saved configural model 
    saved_configural = results.invariant$model_configural,
    # dataframe
    data = df.invariant,
    # model syntax
    model = model.overall, 
    # group variable column in dataframe
    group = "group",
    # number of bootstraps
    nboot = 1000, 
    # which fit index you would like to use
    invariance_index = "cfi",
    # what is your criterion for that fit index
    invariance_rule = .01,
    # what equality constraints are you testing 
    group.equal = c("loadings", "intercepts", "residuals")
  )
```

@@

@@

# examine the loadings 
partial.rs.2$model_coef %>% 
  filter(term == "RS6 ~1") %>% 
  filter(model == "intercepts") %>% 
  dplyr::select(term, group, estimate, std.error)
# examine the fit indices 
partial.rs.2$model_fit %>% 
  filter(model == "intercepts") %>% 
  dplyr::select(AIC, BIC, cfi, tli, rmsea, srmr)

@@

@@

@@

```{r rs14-chinese-pic, include = T, fig.cap = "The differences in intercepts for the second item on the RS14 by language sample. The differences between intercepts are shown on the left hand side with a clear separation between lines. The latent means also show a clear difference between groups where the English group appears to have lower scores overall than the Chinese group."}
load("manu_data/fig-chinese-rs2.Rdata")
complete
```

>>>>

Therefore, purpose of this manuscript is to describe an *R* package, `visualizemi`, that provides functionality to calculate multi-group confirmatory factor analysis, partial invariance tests, visualizations of the size of non-invariance, and potential effect sizes for overall models and individual parameters. No known visualization techniques have been proposed for measurement invariance. By creating panel visualizations, we can supplement a researcher's ability to judge the strength of the non-invariance differences and effect size for each item. The proposed effect sizes demonstrate the likelihood of replication with a similar sample as compared to a randomly assigned group model, thus, illustrating what type of measurement one might expect to find, and how different that is from random chance. Within this technique, the individual parameter effect sizes can be calculated: both the group differences within a model as compared to random and the likelihood of a parameter replication compared to random groups. Coupled with other indicators (i.e., fit indices differences, $d_{MACS}$), we can move toward a better understanding of how much measurement non-invariance is meaningful. This tutorial and package will help researchers plan future studies and aid in the ability to estimate a smallest effect of interest for measurement invariance studies, rather than relying on fit indices and rules of thumb alone.

@@

1.  Learn how to use *visualizemi* to analyze multi-group confirmatory factor analysis, examine partial for invariance, and create visualizations of parameters.
2.  Learn how to estimate the potential replication of multi-group models and their parameters using bootstrapping compared to a random group model.
3.  Be able to calculate and interpret effect sizes for model and parameter replication, as well as parameter group differences.
4.  Understand the impact of measurement variability on replication and generalizability.

@@

1)  `complete`: The output from this model can be found in Figure \@ref(fig:invariant-pic). On the left-hand side, the item invariance is plotted, and on the right-hand side, the latent mean distributions for the two groups are plotted. In the item invariance sub-plot, the visualization includes all three components traditionally seen in MGCFA testing steps: loadings, intercepts, and residuals. Each visualization element was designed to match the traditional visualization for that type of output. All parameter estimates are plotted on the unstandardized estimates and their confidence interval based on the standard error of the estimate. All plots are made with *ggplot2* and *cowplot* [@wilke2020].

@@

2)  `intercept`: Only the left-hand side of the complete plot designed to represent intercepts and factor loadings. Factor loadings represent the slope of the regression equation for the latent variable predicting the scores on the observed variable ($\hat{Y} \sim b_0 + b_1X + \epsilon$). The y-axis indicates the observed variable scores, and here, the plot includes the entire range of the scale of the data for item four. The coefficient ($b_1$) for group 1 was `r apa_num(results.invariant$model_coef %>% filter(model == "Configural") %>% filter(grepl("q4", term)) %>% filter(op == "=~") %>% slice_head() %>% pull(estimate))`, while the coefficient for group 2 was `r apa_num(results.invariant$model_coef %>% filter(model == "Configural") %>% filter(grepl("q4", term)) %>% filter(op == "=~") %>% slice_tail() %>% pull(estimate))`. The ribbon bands around the plotted slopes indicate the confidence interval for that estimate. In this plot, while the coefficients for each group are not literally equal, the overlapping and parallel slope bands indicate they are not different practically.

@@

3)  `mean`: The right-hand side of the complete plot graphing the latent variable means and density from the data. The latent variable is shown on the x-axis using standardized values (i.e., *z*-scores) where -1 indicates one standard deviation below the mean for the latent variable, 0 indicates the mean for the latent variable and so on. The lines indicate the means of the latent variables from the simulated dataset. Group labels are represented in the figure caption on the bottom. Group 1 is usually the group that is alphabetically first in the data set or whichever group is the first that appears when using the `levels()` command.

@@

```{r boot-rr, echo = T, eval = F, size = "small"}
boot.model.invariant <- 
  bootstrap_model(
    # saved configural model 
    saved_configural = results.invariant$model_configural,
    # dataframe
    data = df.invariant,
    # model syntax
    model = model.overall, 
    # group variable column in dataframe
    group = "group",
    # number of bootstraps
    nboot = 1000, 
    # which fit index you would like to use
    invariance_index = "cfi",
    # what is your criterion for that fit index
    invariance_rule = .01,
    # what equality constraints are you testing 
    group.equal = c("loadings", "intercepts", "resduals")
  )
```

@@

```{r echo = F, eval = F}
saveRDS(boot.partial.invariant, "manu_data/boot.partial.invariant.RData")
```

@@

# examine the loadings 
partial.rs.2$model_coef %>% 
  filter(term == "RS6 ~1 ") %>% 
  filter(model == "intercepts") %>% 
  dplyr::select(term, group, estimate, std.error)
# examine the fit indices 
partial.rs.2$model_fit %>% 
  filter(model == "intercepts") %>% 
  dplyr::select(AIC, BIC, cfi, tli, rmsea, srmr)

@@

```{r}
# saveRDS(boot.partial.rs, "manu_data/boot.partial.rs.RData")
boot.partial.rs <- readRDS("manu_data/boot.partial.rs.RData")
```

@@

get_legend <- function(plot, legend = NULL) {

    gt <- ggplotGrob(plot)

    pattern <- "guide-box"
    if (!is.null(legend)) {
      pattern <- paste0(pattern, "-", legend)
    }

    indices <- grep(pattern, gt$layout$name)

    not_empty <- !vapply(
      gt$grobs[indices],
      inherits, what = "zeroGrob",
      FUN.VALUE = logical(1)
    )
    indices <- indices[not_empty]

    if (length(indices) > 0) {
      return(gt$grobs[[indices[1]]])
    }
    return(NULL)
  }

@@

```{r rs14-chinese-pic, include = T, fig.cap = "The differences in intercepts for the second item on the RS14 by language sample. The differences between intercepts are shown on the left-hand side with a clear separation between lines. The latent means also show a clear difference between groups where the English group appears to have lower scores overall than the Chinese group."}
load("manu_data/fig-chinese-rs2.Rdata")
complete
```
